Type theory in type theory using single substitutions

### 2023-12-27 main.tex
stuff

\begin{abstract}
Type theory can be described as a generalised algebraic theory which
automatically gives a notion of model and the existence of the syntax
as the initial model given by a quotient inductive-inductive
type. Such descriptions include categories with families (CwF),
contextual categories, Awodey's natural models, C-systems and
B-systems. With the exception of B-systems, these theories are based on
a parallel substitution calculus where substitutions form a
category. In this paper we define a single substitution calculus (SSC)
for type theory and show that the SSC syntax and the CwF syntax are
isomorphic for a theory with dependent function space and universes.
SSC only includes single substitutions and single weakenings, and six
equations relating these. Four equations describe how to substitute
variables, and two more equations are needed to typecheck these four
equations. SSC provides a simple, minimalistic alternative to parallel
substitution calculi or B-systems for defining type theory. All the
results were formalised in Agda.
\end{abstract}

>>>>

\begin{abstract}
TODO: rewrite the abstract to start with ``tutorial intro''.

Type theory can be described as a generalised algebraic theory which
automatically gives a notion of model and the existence of the syntax
as the initial model given by a quotient inductive-inductive
type. Such descriptions include categories with families (CwF),
contextual categories, Awodey's natural models, C-systems and
B-systems. With the exception of B-systems, these theories are based on
a parallel substitution calculus where substitutions form a
category. In this paper we define a single substitution calculus (SSC)
for type theory and show that the SSC syntax and the CwF syntax are
isomorphic for a theory with dependent function space and universes.
SSC only includes single substitutions and single weakenings, and six
equations relating these. Four equations describe how to substitute
variables, and two more equations are needed to typecheck these four
equations. SSC provides a simple, minimalistic alternative to parallel
substitution calculi or B-systems for defining type theory. All the
results were formalised in Agda.
\end{abstract}
